---
title: "Biostatistical Methods I - HW1"
author: "JingYao Geng"

date: "9/24/2020"
output: github_document
---
```{r setup, message = FALSE, warning = FALSE, echo = FALSE}
# Clean all objects from the current workspace (R memory) 
rm(list = ls())

# load library
library(arsenal)
library(dplyr)
library(ggplot2)
library(tidyverse)
```

## Problem 1

```{r import data, message = FALSE, warning = FALSE, echo = FALSE}
# Import data
Antibodies = read_csv(file = "./Antibodies.csv")
```

#### 1.1.a: Descriptive Statistics

* The **Antibodies** data set contains a total of **1491** observations and **5** variables. 
  * **1224** missing values are founded in variable called **Antibody_IgM**. 
  * **34** observations of variable **Smell** are categorized as  "Unanswered/others". 
  * The rest of variables: **Subject**, **AgeCategory**,  and **Gender**  contain no missing values.

* The following table shows the **descriptive statistics** for all the variable of interest.
  * **Numeric variable: Subject, Antibody_IgM**
  * **Categorical variable: AgeCategory, Gender, Smell**

```{r descriptive statistics, message = FALSE, collapse = FALSE, warning = FALSE, echo = FALSE}
# Change variable names/labels
my_labels <- list(AgeCategory = "Age", Antibody_IgM = "Antibody(IgM)", Smell = "Smell", Gender = "Gender")

# Clean the output
my_controls <- tableby.control(
  total = T,
  test = F,  # No test p-values yet
  numeric.stats = c("meansd", "medianq1q3", "range", "Nmiss2"),
  cat.stats = c("countpct", "Nmiss2"), 
  stats.labels = list(
    meansd = "Mean (SD)",
    medianq1q3 = "Median (Q1, Q3)",
    range = "Min - Max",
    Nmiss2 = "Missing",
    countpct = "N (%)"))

tab1 <- tableby( ~ Subject + Antibody_IgM + AgeCategory + Smell + Gender, data = Antibodies, control = my_controls)

#Descriptive Statistics:
summary(tab1, title = "Descriptive Statistics: Antibodies Data", text = T, labelTranslations = my_labels)


```



#### 1.1.b:  Histogram for the Ig_M values

The histogram of Antibody_IgM shows a right skewed distribution, where the mean of the Antibodies_IgM level tends to be 'pulled' toward the right tail of the distribution.

```{r choose bars, message = FALSE, collapse = TRUE, warning = FALSE, echo = FALSE}
# Choose the number of bars/breaks to use for the histogram

#min(Antibodies$Antibody_IgM, na.rm = TRUE) # 0.048
#max(Antibodies$Antibody_IgM, na.rm = TRUE) #1.0475

```

```{r histgram-IgM, message = FALSE, warning = FALSE, echo = FALSE}
# the shape of Ig_M values looks like a reight-skewd distribution
ggplot(Antibodies, aes(x = Antibody_IgM)) +
  geom_histogram(col = "black", fill = "pink", breaks = seq(0.00,1.20, 0.02)) +                      
  #geom_vline(aes(xintercept = mean(Antibodies$Antibodies_IgM, na.rm = T)), color = "red", linetype = "dashed", size = 2) +  
  labs(title = "Histogram of Antibodies_IgM Level") +                                              
  labs(x = "Antibody IgM", y = "Count")
```

#### 1.2.a:  Descriptive Statistics Stratified by Smell

* There are more observations in the Altered collection of Smell, regardless the variable Age, Gender, and Antibody_IgM.
* The normal group of Semll is more evenly distributed in gender than the altered group.
* Age range 31-50 contributes more than 50% of the observations in both normal and altered groups.

```{r stratified by smell, message = FALSE, collapse = TRUE, warning = FALSE, echo = FALSE}
# filter variable 'Smell' with outcomes "Altered" and "Normal"
Antibodies_Smell = filter(Antibodies, Smell == c("Altered", "Normal"))

# Tabulation by Smell categories
tab2 = tableby( Smell ~ AgeCategory + Gender + Antibody_IgM, data = Antibodies_Smell, control = my_controls)
summary(tab2, title = "Descriptive Statistics Stratified by Smell", labelTranslations = my_labels, text = T, total = F)

```


#### 1.2.b:  Group Boxplots and Side-by-side Histogram

**Boxplots**:

```{r boxplots, echo = FALSE, message = FALSE, warning = FALSE}
# A grouped boxplot
ggplot(Antibodies_Smell, aes(x = Antibody_IgM, fill = Smell)) +
  geom_boxplot(outlier.colour = "black", outlier.shape = 20,
               outlier.size = 1, notch = F) +
  scale_fill_manual(values = c("blue", "red"),            # Change colors
                    labels = expression("Normal", "Altered")) +     # Edit legend labels       
  xlab("Antibodies_IgM Level") +
  ylab("Smell Category") +
  labs(title = "Antibody (IgM) Levels based on Smell")
  theme(legend.position = "right", legend.title = element_blank())
  
```

* Both normal and altered boxes contain missing values.
* The mean of normal group is slightly higher than the altered group.
* The range of both groups are overlapped a lot. 

**Histogram**:

```{r side by side histgrams, message = FALSE, warning = FALSE, echo = FALSE}
# Overlaid histograms
# Histogram of Igm by smell categories 'Normal' and 'Altered'.

ggplot(Antibodies_Smell, aes(x = Antibodies_Smell$Antibody_IgM, fill = factor(Smell))) +
  geom_histogram(binwidth = .5, alpha = .5, position = "identity", col = "black", breaks = seq(0.00,1.10,0.01)) +
  geom_vline(aes(xintercept = mean(Antibodies_Smell$Antibody_IgM, na.rm = T)), color = "red", linetype = "dashed", size = 0.5) +                   
  labs(title = "Histogram of Antibody IgM level Categorized by Smell ") +                                             
  labs(x = "Antibody_IgM Level", y = "Count") + 
  scale_fill_discrete(name = "Smell Category", labels = c("Altered", "Normal"))  

```

* Both Altered and Normal groups display a right skewed distribution overall. With the increase of Antibody_IgM level, the number of observations decrease for both groups.
* There are more observations in the altered group.

